---
title: "Session 6 lab exercise: Survival Analysis part 1"
author: "Levi Waldron"
institute: "CUNY SPH Biostatistics 2"
clean: false
output:
  html_document:
    toc: yes
    df_print: paged
    theme: lumen
    number_sections: yes
  md_document:
    preserve_yaml: false
always_allow_html: true
---

**Learning objectives**

1. Calculate Kaplan-Meier estimates of survival probability over time
2. Plot survival curves for censored time-to-event data
3. Perform and interpret log-rank test
4. Define "informative" censoring

**Exercises**

1. Calculate the follow-up table for 6 MP patients in the leukemia study

```{r}
library(readr)
leuk <- read_csv("leuk.csv")
```

2. Plot the Kaplan-Meier estimate of the follow-up table from 1. `library(survminer)` is recommendable. 

```{r}
library(survival)
library(survminer)
with(leuk, Surv(time, cens))
fit <- survminer::surv_fit(Surv(time, cens) ~ group, data = leuk)
survminer::ggsurvplot(fit,
           xlab = "Time (weeks)",
           ylab = "Survival Probability",
           risk.table = TRUE)
```


3. What is the 75th percentile of survival times for the 6 MP group? For the Placebo group? This is the time that 75% of the patients survive. This is also the time at which 25% of patients have had events.

```{r}
quantile(fit)
```

```{r}
survdiff(Surv(time, cens) ~ group, data = leuk)
```

4. Suppose you were instructed to cap follow-up times at 20 weeks. Re-do the Kaplan-Meier plot for both groups, and re-do the logrank test.

```{r}
library(tidyverse)
leuknew <- leuk %>% mutate(newtime = pmin(time, 20)) %>% 
  mutate(newcens = ifelse(time <= 20, cens, 0))
```

Kaplan-Meier plot
```{r}
fit <- survminer::surv_fit(Surv(newtime, newcens) ~ group, data = leuknew)
survminer::ggsurvplot(fit,
           xlab = "Time (weeks)",
           ylab = "Survival Probability",
           risk.table = TRUE)
```

Logrank test
```{r logrank}
survdiff(Surv(newtime, newcens) ~ group, data = leuknew)
```


5. Give a hypothetical example of how censoring in this example might be "informative."

If sicker patients were moved to another hospital where they weren't followed up on, and were also more likely to relapse, this would be informative censoring.

